---
title: "Analysis on Differences"
author: "James Wu"
date: "7/7/2020"
output: html_document
---

```{r setup, include=FALSE}
load("./Comorbidity_df.RData")
library(data.table)
library(stringr)
library(leaflet)
library(sf)
library(htmltools)
library(dplyr)
```

## Introduction
After creating the Severity Index (SI), I wanted to see how well they matched up with the actual fatality data and their locations. New York City Department of Health and Mental Hygiene publishes their fatality numbers within modified zip codes (for further information on how they aggregated deaths, please refer to their Github page: https://github.com/nychealth/coronavirus-data).

## Data Preparation
Since the fatality data is given at the modified zip code level, the SIs need to be crosswalked up to that level for a proper comparison. We assume that the SI we created are uniform within each census tract, so that we can extrapolate the SI of a modified zip as a weighted average of SIs of the census tracts that make up the modified zip. This is done by multiplying each census tract’s SI by their population, summed at the modified zip level, and then divided by the modified zip’s total population.

```{r}
# Crosswalk between census tract and zcta
URL_ZCTAtoCT <- "https://www2.census.gov/geo/docs/maps-data/data/rel/zcta_tract_rel_10.txt?#"
ZCTAtoCT_df <- fread(URL_ZCTAtoCT)
ZNYC <- ZCTAtoCT_df[STATE=="36" & (COUNTY=="05" | COUNTY=="5" | COUNTY=="47" | COUNTY=="81" | COUNTY=="85" | COUNTY=="61")]

# Get fatality data from DOHMH
URL_Fatality<- "https://raw.githubusercontent.com/nychealth/coronavirus-data/master/data-by-modzcta.csv"
Fatality_df <- fread(URL_Fatality)

# Get MZCTA conversion data from DOHMH
URL_ZCTAtoMZCTA <- "https://raw.githubusercontent.com/nychealth/coronavirus-data/master/Geography-resources/ZCTA-to-MODZCTA.csv"
ZCTAtoMZCTA_df <- fread(URL_ZCTAtoMZCTA)

# Merge Fatality Data and MODZCTA with ZNYC
ZNYC[,"MODIFIED_ZCTA" := .(sapply(ZCTA5, function(x) ZCTAtoMZCTA_df[ZCTA == x,MODZCTA]))]
ZNYC_Fatality <- merge(ZNYC, Fatality_df,all.x = TRUE)

# Convert boroughs to the ones we have
ZNYC_Fatality[,boro_code := stringr::str_sub(GEOID,-8,-7)] #Extract the boro code from GEOID
ZNYC_Fatality[,TRACT := as.character(TRACT/100)]
ZNYC_Fatality[TRACT == "170.1",TRACT := "170.10"] #Census Tract 170.1 is coded as '170.10' in our shapefile

Convertboro <- data.table(Census_boro_code = c('05','47','81','85','61'), 
                          boro_code=c(2,3,4,5,1), 
                          boro_name = c("Bronx","Brooklyn","Queens","Staten Island","Manhattan"))

ZNYC_Fatality[,c("boro_code","boro_name") := .(sapply(boro_code, function(x) Convertboro[Census_boro_code == x,boro_code]),
                                               sapply(boro_code, function(x) Convertboro[Census_boro_code == x,boro_name]))][
                                                 ,name := paste0('Census Tract ',TRACT,', ',boro_name,' Boro')]

# Create the percentage variables
## County and Tract Population
ZNYC_Fatality[, `:=` (COUNTYTRACTPOP = sum(POPPT)), by = c("COUNTY","TRACT")]
### Percentage of Population in Census Tract accounted for in the County
ZNYC_Fatality[, `:=` (PERCENT_COUNTYTRACT = POPPT/COUNTYTRACTPOP)]

## Zip and County Population
### ZCTA5 Population
ZNYC_Fatality[, `:=` (ZCTAPOP = sum(POPPT)), by = "ZCTA5"]

## Modified Zip and Zip Population
### Modified Zip Population
ZNYC_Fatality[, `:=` (MODIFIEDZCTAPOP = sum(POPPT)), by = "MODIFIED_ZCTA"]
### Population of Zip / Modified Zip
ZNYC_Fatality[, PERCENT_MODZZCTA := (ZCTAPOP)/MODIFIEDZCTAPOP]

# Merge severity index variable with Zip and Fatality data
ZNYC_Fatality <- merge(x = ZNYC_Fatality, y = Comorbidities_df[ , c("name", "severity_index","severity_index_norm")], by = "name", all.x=TRUE)

# Create weighted SI by MZCTA
## Multiply SI with POPPT and COUNTYTRACTPOP
ZNYC_Fatality[, RAW_MZCTA_SI := severity_index_norm * POPPT]
## Divide by MODIFIEDZCTAPOP and sum them up for a weighted average SI by MODIFIED_ZCTA
ZNYC_Fatality[,`:=` (WEIGHTED_MZCTA_SI = sum(RAW_MZCTA_SI,na.rm=T)/MODIFIEDZCTAPOP),by = "MODIFIED_ZCTA"]
```

## Correlation
First, let's look at how correlated the new SIs at the modified zip code level and the fatality rates are. Plot the newly formed SIs against the death rates provided by NYCDOHMH:
```{r}
# Plot it!
library(plotly)
MZCTA_df <- unique(ZNYC_Fatality, by = c("COVID_DEATH_RATE","WEIGHTED_MZCTA_SI"))


## SI vs. COVID Death Rate
plotly_labels <- paste("<b>",paste0("Modified ZIP: ",MZCTA_df$MODIFIED_ZCTA),"</b><br>",
                       "<b>Neighborhood:",MZCTA_df$NEIGHBORHOOD_NAME,"</b><br>",
                       "<b>SI:</b>",round(MZCTA_df$WEIGHTED_MZCTA_SI,2),"<br>",
                       "<b>DEATH RATE:</b>",MZCTA_df$COVID_DEATH_RATE)


fig <- plot_ly(data = MZCTA_df,
               x = ~WEIGHTED_MZCTA_SI,
               y = ~COVID_DEATH_RATE,
               type = 'scatter',
               mode = 'markers',
               name = "",
               hoverinfo = 'text',
               text = lapply(plotly_labels,HTML)) %>%
  layout(
    title = "Severity Index vs. NYC Death Rates",
    xaxis = list(title = "Severity Index"),
    yaxis = list(title = "COVID Death Rate (per 100,000)"))

lm_fit <- lm(COVID_DEATH_RATE~WEIGHTED_MZCTA_SI, data = MZCTA_df)

fig %>%
  add_trace(x = ~WEIGHTED_MZCTA_SI, y = fitted(lm_fit), mode = "lines") %>%
  layout(showlegend = F)
```

Here, we see a general upwards trend, a positive correlation between our created severity indices and the area’s death rates (0.67), which could be interpreted as the area’s comorbidity makeup being correlated with their death rates. The SIs were also able to accurately identify the area with the highest death rates (East New York) and the area with the lowest death rates (Financial District).

## Map of Differences
Since the SIs are a constructed index and do not have a scale, it may be more interesting to see how well the relative quantiles for the SI match up with the relative quantiles for the COVID Death Rates. The following map shows this:
```{r, message=F}
# Map of areas with highest fatalities
URL_MZCTA <- "./MZCTA_2010"
MZCTA_shp <- st_read(URL_MZCTA) %>%
  st_transform("+proj=longlat +datum=WGS84")

MZCTA_sf <- st_sf(merge(MZCTA_df[,MODIFIED_ZCTA := as.factor(MODIFIED_ZCTA)],
                        MZCTA_shp,
                        by.x = "MODIFIED_ZCTA",
                        by.y = "MODZCTA"))

# Consider breaking the SIs and Death Rates into 9 bins, how well do the SIs match up quantile-wise?
bins_MODSI <- round(c(0,quantile(MZCTA_sf$WEIGHTED_MZCTA_SI, 0.15, quantile(MZCTA_sf$WEIGHTED_MZCTA_SI, 0.30, na.rm = T),na.rm = T),quantile(MZCTA_sf$WEIGHTED_MZCTA_SI, 0.40, na.rm = T),quantile(MZCTA_sf$WEIGHTED_MZCTA_SI, 0.50, na.rm = T),quantile(MZCTA_sf$WEIGHTED_MZCTA_SI, 0.60, na.rm = T),quantile(MZCTA_sf$WEIGHTED_MZCTA_SI, 0.70, na.rm = T),quantile(MZCTA_sf$WEIGHTED_MZCTA_SI, 0.80, na.rm = T),quantile(MZCTA_sf$WEIGHTED_MZCTA_SI, 0.90, na.rm = T),quantile(MZCTA_sf$WEIGHTED_MZCTA_SI, 0.99, na.rm = T), max(MZCTA_sf$WEIGHTED_MZCTA_SI, na.rm = T) + 1),4)
bins_COVIDDEATH <- round(c(0,quantile(MZCTA_sf$COVID_DEATH_RATE, 0.15, quantile(MZCTA_sf$COVID_DEATH_RATE, 0.30, na.rm = T),na.rm = T),quantile(MZCTA_sf$COVID_DEATH_RATE, 0.40, na.rm = T),quantile(MZCTA_sf$COVID_DEATH_RATE, 0.50, na.rm = T),quantile(MZCTA_sf$COVID_DEATH_RATE, 0.60, na.rm = T),quantile(MZCTA_sf$COVID_DEATH_RATE, 0.70, na.rm = T),quantile(MZCTA_sf$COVID_DEATH_RATE, 0.80, na.rm = T),quantile(MZCTA_sf$COVID_DEATH_RATE, 0.90, na.rm = T),quantile(MZCTA_sf$COVID_DEATH_RATE, 0.99, na.rm = T), max(MZCTA_sf$COVID_DEATH_RATE, na.rm = T) + 1),4)

cuts_MODSI <- cut(MZCTA_sf$WEIGHTED_MZCTA_SI,
                  bins_MODSI, 
                  labels = c(1,2,3,4,5,6,7,8,9))
cuts_COVIDDEATH <- cut(MZCTA_sf$COVID_DEATH_RATE,
                       bins_COVIDDEATH, 
                       include.lowest=TRUE,
                       labels = c(1,2,3,4,5,6,7,8,9))

cuts_diff <- abs(as.numeric(cuts_MODSI) - as.numeric(cuts_COVIDDEATH)) + 1

# Create Labels and color bins for map
labels_MOD<- paste("<h4>",paste0("Neighborhood Name: ",MZCTA_sf$NEIGHBORHOOD_NAME),"</h4>",
                   "<h4>","Modified Zip: ",MZCTA_sf$MODIFIED_ZCTA,"</h4>",
                   "<p>","Modified SI: ",format(MZCTA_sf$WEIGHTED_MZCTA_SI,nsmall = 0, big.mark = ","),"</p>",
                   "<p>","COVID Death Rate (per 100,000): ",format(MZCTA_sf$COVID_DEATH_RATE,nsmall = 0,big.mark = ","),"</p>")

labels_Diff<- paste("<h4>",paste0("Neighborhood Name: ",MZCTA_sf$NEIGHBORHOOD_NAME),"</h4>",
                    "<h4>","Modified Zip: ",MZCTA_sf$MODIFIED_ZCTA,"</h4>",
                    "<p>","Modified SI: ",format(MZCTA_sf$WEIGHTED_MZCTA_SI,nsmall = 0, big.mark = ","),"</p>",
                    "<p>","COVID Death Rate (per 100,000): ",format(MZCTA_sf$COVID_DEATH_RATE,nsmall = 0,big.mark = ","),"</p>",
                    "<p>","Quantile Difference: ",format(cuts_diff,nsmall = 0, big.mark = ","),"</p>")

pal_MODSI <- colorBin("YlOrRd", domain = MZCTA_sf$WEIGHTED_MZCTA_SI, bins = bins_MODSI)
pal_COVIDDEATH <- colorBin("YlOrRd", domain = MZCTA_sf$COVID_DEATH_RATE, bins = bins_COVIDDEATH)
pal_diff <- colorNumeric("RdPu", domain = cuts_diff)

# Map it!
leaflet(MZCTA_shp, options = leafletOptions(zoomSnap = 0.5, 
                                            zoomDelta = 0.5)) %>%
  setView(-73.935242,40.730610,10.5) %>%
  addProviderTiles("CartoDB.Positron") %>%
  #Polygon Layers
  addPolygons(data = MZCTA_sf,
              weight = 1,
              color = "grey",
              group = "Differences",
              fillOpacity = 1,
              fillColor = pal_diff(cuts_diff),
              label = lapply(labels_Diff,HTML)
  ) %>%
  addPolygons(data = MZCTA_sf,
              weight = 1,
              color = "grey",
              group = "Modified SI",
              fillColor = pal_MODSI(MZCTA_sf$WEIGHTED_MZCTA_SI),
              label = lapply(labels_MOD,HTML)) %>%
  addPolygons(data = MZCTA_sf,
              weight = 1,
              color = "grey",
              group = "COVID19 Death Rate",
              fillColor = pal_COVIDDEATH(MZCTA_sf$COVID_DEATH_RATE),
              label = lapply(labels_MOD,HTML)) %>%
  #Legend Layers
  addLegend("bottomleft", pal = pal_diff, values = ~cuts_diff,
            group = "Differences",
            title = "Accuracy",
            opacity = 0.7,
            labFormat =  function(type, cuts, p) {
              n = length(cuts) 	
              cuts[n] = "way off" 	
              for (i in 2:(n-1)){cuts[i] = ""} 	
              cuts[1] = "on point" 	
              paste0(cuts[-n], cuts[-1])}) %>%
  addLegend("bottomright", pal = pal_MODSI, values = ~MZCTA_sf$WEIGHTED_MZCTA_SI,
            group = "Modified SI",
            title = "Severity",
            opacity = 0.7,
            labFormat =  function(type, cuts, p) {
              n = length(cuts) 	
              cuts[n] = "High" 	
              for (i in 2:(n-1)){cuts[i] = ""} 	
              cuts[1] = "Low" 	
              paste0(cuts[-n], cuts[-1])}) %>%
  #Layers Control
  addLayersControl(
    baseGroups =c("COVID19 Death Rate", "Modified SI","Differences"),
    options = layersControlOptions(collapsed=FALSE),
    position = "topright"
  )
```

## Conclusions
The SIs were able to accurately identify the severity level of many locations. However, there are certain areas that are significantly different than expected (i.e. large quantile difference between SIs and COVID Death Rates); areas like Corona, Elmhurst, Jackson Heights, Kew Gardens, and Battery Park City have much higher death rates than what the SIs would suggest, and areas like Roosevelt Island, Wakefield/Woodlawn, Hunts Point, and Breezy Point have much lower death rates than what the SIs would suggest. This may imply that there are other driving forces and confounding variables that must be accounted for, and just the comorbidity makeup is not enough for these areas. Further research in these areas and a more rigorous approach in our assumptions may be warranted for future studies. 